Exploration

### 2018-03-13 exploration.Rmd
Nt as vector instead of matrix

#---- Truth: use assigned slopes to fill IPM matrix
U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
                       p, n_z, n_x, X, sdd.pr, env.df)

#---- Realization: generate simulated data
S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
                   p, X, n_z, sdd, U$sdd.j)

#---- Aggregate results
lam.df <- env.df
lam.df$lambda <- apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S=S$N_sim[,tmax]/(S$N_sim[,tmax-1]+.01),
             nSeed=S$nSd[,tmax], D=S$D[,tmax], 
             B0=S$B[,1], Btmax=S$B[,tmax+1], 
             N.S=map_dbl(S$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S=map_dbl(S$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S=map_dbl(S$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
  add_column(N.U=apply(U$Nt[,,,tmax],3,sum), lam.U=U$lam.t[,tmax-1])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.U))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.S>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>1 & lam.U<1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0 & lambda<1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Surv.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Rcr.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
par(mfrow=c(2,2))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
matplot(t(log(apply(U$Nt, 3:4, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
matplot(t(S$N_sim[,2:tmax]/S$N_sim[,1:(tmax-1)]), 
        type="l", lty=1, col=rgb(0,0,0,.1))
matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
#matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))

@@

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Mx=500, CA=20, IPM=10) # number of cells in sample
O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
```

@@

########
## predict distribution with fit models
########
#---- Truth: use estimated slopes to fill IPM matrix
U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
                         p.ipm, n_z, n_x, X, sdd, env.df)

#---- Realization: use estimated slopes to generate simulated data
S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
                     p.ipm, X, n_z, sdd, U.f$sdd.j)

#---- Aggregate results
lam.df$lambda.f <- apply(U.f$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S.f=S.f$N_sim[,tmax]/(S.f$N_sim[,tmax-1]+.01),
             nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
             B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
             N.S.f=map_dbl(S.f$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S.f=map_dbl(S.f$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S.f=map_dbl(S.f$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
  add_column(N.U.f=apply(U.f$Nt[,,,tmax],3,sum), lam.U.f=U.f$lam.t[,tmax-1])

>>>>

#---- Truth: use assigned slopes to fill IPM matrix
U <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.75, discrete=1,
                       p, n_z, n_x, X, sdd.pr, env.df, verbose=T)

#---- Realization: generate simulated data
S <- simulate_data(n.cell, tmax, n_i, z.rng, U$lo, U$hi, 
                   p, X, n_z, sdd, U$sdd.j, verbose=T)

#---- Aggregate results
lam.df <- env.df
lam.df$lambda <- apply(U$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S=S$N_sim[,tmax]/(S$N_sim[,tmax-1]+.01),
             nSeed=S$nSd[,tmax], D=S$D[,tmax], 
             B0=S$B[,1], Btmax=S$B[,tmax+1], 
             N.S=map_dbl(S$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S=map_dbl(S$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S=map_dbl(S$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay=nSeed*(1-p$p_emig), nSdLeave=nSeed*p$p_emig) %>%
  add_column(N.U=apply(U$Nt[-1,,tmax+1],2,sum), lam.U=U$lam.t[,tmax])

########
## plots
########
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(lam.U))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=lambda>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.S>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.S>0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=lam.U>1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=N.U>0 & lam.U<1)) + geom_tile() 
ggplot(lam.df, aes(x=lon, y=lat, fill=Surv.S>0 & lambda<1)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.U))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(N.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Surv.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=log(Rcr.S))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=lon, y=lat, fill=nSeed)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=D + nSdStay)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=Btmax>B0)) + geom_tile()
ggplot(lam.df, aes(x=lon, y=lat, fill=nSdStay > D)) + geom_tile()
par(mfrow=c(2,2))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)
matplot(t(log(U$lam.t)), type="l", lty=1, col=rgb(0,0,0,0.1))
matplot(t(log(apply(U$Nt[-1,,], 2:3, sum))), type="l", lty=1, col=rgb(0,0,0,0.1))
matplot(t(log(S$N_sim[,2:tmax]/(S$N_sim[,1:(tmax-1)]))), 
        type="l", lty=1, col=rgb(0,0,0,.1))
matplot(t(log(S$N_sim+1)), type="l", lty=1, col=rgb(0,0,0,0.1))
#matplot(t(log(nSd*(1-p$p_emig)+D)), type="l", lty=1, col=rgb(0,0,0,0.1))

@@

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
```

@@

########
## predict distribution with fit models
########
#---- Truth: use estimated slopes to fill IPM matrix
U.f <- fill_IPM_matrices(n.cell, tmax, n_i, n, z.rng, buffer=0.5, discrete=1,
                         p.ipm, n_z, n_x, X, sdd, env.df, verbose=T)

#---- Realization: use estimated slopes to generate simulated data
S.f <- simulate_data(n.cell, tmax, n_i, z.rng, U.f$lo, U.f$hi, 
                     p.ipm, X, n_z, sdd, U.f$sdd.j, verbose=T)

#---- Aggregate results
lam.df$lambda.f <- apply(U.f$IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% 
  add_column(lam.S.f=S.f$N_sim[,tmax]/(S.f$N_sim[,tmax-1]+.01),
             nSeed.f=S.f$nSd[,tmax], D.f=S.f$D[,tmax], 
             B0.f=S.f$B[,1], Btmax.f=S.f$B[,tmax+1], 
             N.S.f=map_dbl(S.f$d, ~sum(!is.na(.$sizeNext[.$yr==tmax]))),
             Surv.S.f=map_dbl(S.f$d, ~sum(.$surv[.$yr==tmax], na.rm=T)),
             Rcr.S.f=map_dbl(S.f$d, ~sum(is.na(.$size[.$yr==tmax])))) %>%
  mutate(nSdStay.f=nSeed.f*(1-p.ipm$p_emig), nSdLeave.f=nSeed.f*p.ipm$p_emig) %>%
  add_column(N.U.f=apply(U.f$Nt[,,tmax],2,sum), lam.U.f=U.f$lam.t[,tmax-1])
